---
title: "cx"
author: "Christopher Eshleman"
date: "10/20/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#CX Correlation

Note: While trying to read a SPSS file with the read.spss function I found an error. It was due to label (not variable name) repetition in the raw data. A custom (not from me) function provides a work-around: 

```{r echo=FALSE}
library(foreign) 
Int2Factor <- function(x)
{
  if(!is.null(attr(x, "value.labels"))){
    vlab <- attr(x, "value.labels")
    if(sum(duplicated(vlab)) > 0)
      cat("Duplicated levels:", vlab, "\n")
    else if(sum(duplicated(names(vlab))) > 0)
      cat("Duplicated labels:",
          names(vlab)[duplicated(names(vlab))], "\n")
    else
      x <- factor(x, levels = as.numeric(vlab),
                  labels = names(vlab))
  }
  x
}

```

Now call read.spss with the argument use.value.labels set to FALSE and, later, convert the SPSS categorical variables into R factors with the function: 

```{r echo=FALSE}
cx = read.spss("2013_TxT_CSS_Dep_SPSS_Database.sav", use.value.labels = FALSE) #to.data.frame=TRUE) 
cx = lapply(cx, Int2Factor) 
cx = as.data.frame(cx, stringsAsFactors = FALSE) 


```

A correlation analysis here will need to grapple with the wide (big) nature of the data — nearly 2,600 variables. 